---
title: "Homework 2"
author: "Sarah Vititoe"
date: "9/27/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before beginning, let's load in the tidyverse. 
```{r packages}
library(tidyverse)
```

### Problem 1: NYC Subways 
*"This problem focuses on NYC Transit data; in particular, this CSV file contains information related to each entrance and exit for each subway station in NYC."*

*"Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance." *

*"Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful)."*

```{r}
mta <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, starts_with("station"), starts_with("route"), 
         entry, entrance_type, vending, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

In the code above, I read in my file name using the read_csv function. I piped my result into the janitor package, which helps me clean up the names a bit. Then, I selected the variables we were interested in. We have the line, information about the station (it's name, latitude, and longitude), all of the routes served by the station, a logical variable that tells us whether there is an entrance (as opposed to just an exit), what the entrance type is (e.g. stairs, or terrifying elevator on 168th), and whether than entrance is ADA compliant. I had to coerce the entry variable to be logical, by using an ifelse statement. We have `r ncol(mta)` columns and `r nrow(mta)` rows in our tibble. This dataset is definitely not tidy, since we have a ton of rows with the different routes spread out in various columns. 

*"How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here."*

```{r}
mta %>%
  group_by(line, station_name) %>%
  n_distinct()
```

We have 684 unique stations. 


*"How many stations are ADA compliant?"*
```{r}
mta %>% 
  select(line, station_name, ada) %>%
  group_by(line, station_name) %>%
  filter(ada == TRUE) %>% 
  n_distinct() 
```

Only 84 of those stations are ADA compliant. :(

*"What proportion of station entrances / exits without vending allow entrance?"*"

```{r}
mta %>%
  filter(vending == "NO") %>% 
  select(line, station_name, entry) %>% 
  distinct() %>% 
  summarise(proportion = mean(entry)) %>%
  knitr::kable()
```

38.7% of all station etrances without a vending machine allow entrances. 

*"Reformat data so that route number and route name are distinct variables."*
```{r}
mta_tidy <- mta %>%
  gather(key = route_number, value = route_name, starts_with("route")) %>%
  separate(route_number, into = c("delete", "route_number"), sep = 5) %>%
  select(-delete) %>% 
  select(line, station_name, route_number, route_name, ada, everything()) 
```

Here, we gathered the repetative route variables, and gathered them so that the key was the route number, and the value was the route name. Then, we split the route number so that it would be atomic and contain only the number, and not the word "route" with the number. From there, I dropped my dummy variable that contained the word "route" from splitting it off of route_number, and ordered my dataframe so that important variables would appear first. 

*"How many distinct subway stations serve the A train?"*
```{r}
mta_tidy %>%
  select(line, station_name, route_name, ada) %>% 
  filter(route_name == "A") %>%
  n_distinct()
```

There are 60 distinct stations that serve the A train.

```{r}
mta_tidy %>%
  select(line, station_name, route_name, ada) %>% 
  filter(route_name == "A" & ada == TRUE) %>%
  n_distinct()
```

There are 17 distinct A train stations that are ADA accessible. 

### Problem 2: Mr. Trash Wheel 

Read in and clean Mr.Trash Wheel dataset. Omit columns containing notes, use reasonable variable names, omit rows that do not include dupster-specific data, and round the number of sports_balls.

```{r}
mrtrash <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx",sheet = "Mr. Trash Wheel", range = "A2:N258") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Read in and clean the 2016 and 2017 precipitation data. 
```{r}
precip_16 <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  rename(precip_total = total) %>% 
  mutate(year = "2016") %>% 
  select(year, month, precip_total)

precip_17 <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  rename(precip_total = total) %>% 
  mutate(year = "2017") %>% 
  select(year, month, precip_total)
```

Join and convert numeric month to month name. 
```{r}
months <- tibble(
  abbrv = c(1:12), 
  month_name = month.name
)

precip <- precip_16 %>% 
  union(precip_17) %>% 
  left_join(months, by = c("month" = "abbrv")) %>% 
  select(month_name, year, precip_total) %>% 
  rename(month = month_name)
  
rm(months)
rm(precip_16)
rm(precip_17)
```

In our Mr. Trash Wheel dataframe, we see that we have `r nrow(mrtrash)` observations and `r ncol(mrtrash)` columns. For each dumpster, we can see when the dumpster was run, the volume and weight of the trash it collected, and a count of the types of trash collected. For our precipitation data, we have `r nrow(precip)` months of collected data, and `r ncol(precip)` variables recorderd. We can see the month and year where precipitation were recorded, along with the precipitation total. 

For the year `r precip %>% filter(year == 2017) %>% group_by(year) %>% summarise(sum(precip_total, na.rm = TRUE))` inches of rain were recorded. The median number of sports balls collected per day by Mr. Trash Wheel over the course of 2016 was  `r mrtrash %>% filter(year == 2016) %>% group_by(year) %>%  summarise(median(sports_balls)) %>%  select(-year)`.